---
title: "ALS Downloader session report"
output:
  html_document:
    self_contained: true
params:
  tiles: NULL
  aoi_area_km2: NA
  aoi: NULL
---

<style>
body{max-width:900px;margin:0 auto;padding:0 24px 70px;font-family:-apple-system,"Segoe UI",Helvetica,Arial,sans-serif;color:#222}
h1.title{color:#1c4a5e}
h2{border-bottom:1px solid #e2e2e2;padding-bottom:4px;margin-top:34px}
.als-mission{color:#555;font-size:15px;font-style:italic;border-left:3px solid #8fb4c9;padding-left:12px;margin:10px 0 22px}
.als-footer-badge{position:fixed;right:18px;bottom:14px;background:#152936;color:#dbe7ee;font-size:11.5px;padding:5px 12px;border-radius:14px;letter-spacing:.3px;box-shadow:0 2px 8px rgba(0,0,0,.25)}
table{font-size:13.5px}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "asis")
tiles <- params$tiles
aoi_area_km2 <- params$aoi_area_km2
aoi <- params$aoi
has_tiles <- !is.null(tiles) && is.data.frame(tiles) && nrow(tiles) > 0
has_aoi <- inherits(aoi, "sf") && nrow(aoi) > 0
has_map <- has_tiles && has_aoi && inherits(tiles, "sf")
# The map figure needs geometry; every table below only ever needs the
# plain attribute columns, so it uses this geometry-free copy instead.
tiles_df <- if (has_tiles && inherits(tiles, "sf")) sf::st_drop_geometry(tiles) else tiles
```

<p class="als-mission">Discover, inspect and download airborne LiDAR. Open-source software connecting researchers to aerial laser-scanning point clouds from multiple providers. Draw or upload a study area, inspect acquisition dates and tiles, and download original files without writing code.</p>

Generated `r format(Sys.time(), "%Y-%m-%d %H:%M")` by [ALS Downloader](https://github.com/Cesarito2021/als_downloader). This report summarises a search/download selection at the time it was generated; it does not certify data quality, coverage completeness or licence clearance. Verify each dataset's own terms before use.

## Study area

```{r}
if (length(aoi_area_km2) == 1 && is.finite(aoi_area_km2)) {
  cat(sprintf("Approximately **%.4f km2**.\n\n", aoi_area_km2))
} else {
  cat("Not recorded.\n\n")
}
```

```{r fig.width=7.2, fig.height=5.6, fig.align="center"}
if (has_map) {
  aoi_geo <- sf::st_transform(aoi, 4326)
  tiles_geo <- sf::st_transform(tiles, 4326)
  reported <- ifelse(is.na(tiles_geo$acquired_end), tiles_geo$acquired_start, tiles_geo$acquired_end)
  year <- suppressWarnings(as.integer(substr(reported, 1, 4)))
  years <- sort(unique(stats::na.omit(year)))
  n_year <- length(years)
  cols <- grDevices::hcl.colors(max(n_year, 2), "Viridis")
  fill <- if (n_year > 0) cols[match(year, years)] else rep(NA_character_, nrow(tiles_geo))
  fill[is.na(fill)] <- "#8a97a1"
  combined <- c(sf::st_geometry(aoi_geo), sf::st_geometry(tiles_geo))
  bb <- sf::st_bbox(combined)
  plot(sf::st_geometry(tiles_geo), col = fill, border = "#1c2b35",
    xlim = bb[c("xmin", "xmax")], ylim = bb[c("ymin", "ymax")], main = "")
  plot(sf::st_geometry(aoi_geo), add = TRUE, border = "#dc2626", lwd = 2, col = NA)
  if (n_year > 0) legend("topright", legend = c(as.character(years), "Unknown"),
    fill = c(cols[seq_len(n_year)], "#8a97a1"), bty = "n", cex = .8, title = "Acquisition year")
  cat("\n\n*Tile footprints (fill = acquisition year) and the study area outline (red). Geometry only, not a basemap image.*\n\n")
} else if (has_tiles) {
  cat("Study area geometry was not supplied to this report; only the tile footprints below are known.\n\n")
}
```

## Summary

```{r}
if (!has_tiles) {
  cat("No tiles in this selection.\n")
} else {
  known <- is.finite(tiles_df$size_bytes)
  total_gb <- sum(tiles_df$size_bytes[known], na.rm = TRUE) / 1024^3
  providers <- sort(unique(tiles_df$provider))
  cat(sprintf("- **%d** tile(s) in this selection\n", nrow(tiles_df)))
  cat(sprintf("- **%.3f GB** known total size (%d file(s) with unknown size)\n", total_gb, sum(!known)))
  cat(sprintf("- **%d** provider(s): %s\n", length(providers), paste(providers, collapse = ", ")))
}
```

### By provider

```{r}
if (has_tiles) {
  size <- ifelse(is.finite(tiles_df$size_bytes), tiles_df$size_bytes, 0)
  by_provider <- stats::aggregate(size, by = list(Provider = tiles_df$provider), FUN = sum)
  names(by_provider)[2] <- "known_bytes"
  counts <- as.data.frame(table(Provider = tiles_df$provider), responseName = "Tiles")
  by_provider <- merge(counts, by_provider, by = "Provider")
  by_provider[["Known size (GB)"]] <- round(by_provider$known_bytes / 1024^3, 2)
  by_provider$known_bytes <- NULL
  print(knitr::kable(by_provider))
}
```

## Tiles

```{r}
if (has_tiles) {
  cols <- intersect(c("filename", "dataset", "provider", "acquired_start", "acquired_end", "size_bytes"), names(tiles_df))
  show <- tiles_df[, cols, drop = FALSE]
  if ("size_bytes" %in% names(show)) show$size_bytes <- round(show$size_bytes / 1024^2, 2)
  names(show) <- c("File", "Dataset", "Provider", "Collection start", "Collection end", "Size (MB)")[seq_along(cols)]
  print(knitr::kable(show, row.names = FALSE))
}
```

## Citations and licences

```{r}
if (has_tiles) {
  cols <- intersect(c("citation", "license_url"), names(tiles_df))
  if (length(cols) == 2) {
    pairs <- unique(tiles_df[, cols, drop = FALSE])
    for (i in seq_len(nrow(pairs))) cat("-", pairs$citation[i], "-", pairs$license_url[i], "\n\n")
  }
}
```

---

Collection dates are provider-reported and may be unknown; publication or filename dates are never substituted. Downloads preserve original tiles, including portions outside the study area. This report is generated locally and is not transmitted anywhere.

<div class="als-footer-badge">ALS Downloader</div>
