---
title: "ALS Downloader session report"
output:
  html_document:
    self_contained: true
params:
  tiles: NULL
  aoi_area_km2: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "asis")
tiles <- params$tiles
aoi_area_km2 <- params$aoi_area_km2
has_tiles <- !is.null(tiles) && is.data.frame(tiles) && nrow(tiles) > 0
```

Generated `r format(Sys.time(), "%Y-%m-%d %H:%M")` by [ALS Downloader](https://github.com/Cesarito2021/als_downloader). This report summarises a search/download selection at the time it was generated; it does not certify data quality, coverage completeness or licence clearance. Verify each dataset's own terms before use.

## Study area

```{r}
if (length(aoi_area_km2) == 1 && is.finite(aoi_area_km2)) {
  cat(sprintf("Approximately **%.4f km2**.\n", aoi_area_km2))
} else {
  cat("Not recorded.\n")
}
```

## Summary

```{r}
if (!has_tiles) {
  cat("No tiles in this selection.\n")
} else {
  known <- is.finite(tiles$size_bytes)
  total_gb <- sum(tiles$size_bytes[known], na.rm = TRUE) / 1024^3
  providers <- sort(unique(tiles$provider))
  cat(sprintf("- **%d** tile(s) in this selection\n", nrow(tiles)))
  cat(sprintf("- **%.3f GB** known total size (%d file(s) with unknown size)\n", total_gb, sum(!known)))
  cat(sprintf("- **%d** provider(s): %s\n", length(providers), paste(providers, collapse = ", ")))
}
```

### By provider

```{r}
if (has_tiles) {
  size <- ifelse(is.finite(tiles$size_bytes), tiles$size_bytes, 0)
  by_provider <- stats::aggregate(size, by = list(Provider = tiles$provider), FUN = sum)
  names(by_provider)[2] <- "known_bytes"
  counts <- as.data.frame(table(Provider = tiles$provider), responseName = "Tiles")
  by_provider <- merge(counts, by_provider, by = "Provider")
  by_provider[["Known size (GB)"]] <- round(by_provider$known_bytes / 1024^3, 3)
  by_provider$known_bytes <- NULL
  print(knitr::kable(by_provider))
}
```

## Tiles

```{r}
if (has_tiles) {
  cols <- intersect(c("filename", "dataset", "provider", "acquired_start", "acquired_end", "size_bytes"), names(tiles))
  show <- tiles[, cols, drop = FALSE]
  names(show) <- c("File", "Dataset", "Provider", "Collection start", "Collection end", "Size (bytes)")[seq_along(cols)]
  print(knitr::kable(show, row.names = FALSE))
}
```

## Citations and licences

```{r}
if (has_tiles) {
  cols <- intersect(c("citation", "license_url"), names(tiles))
  if (length(cols) == 2) {
    pairs <- unique(tiles[, cols, drop = FALSE])
    for (i in seq_len(nrow(pairs))) cat("-", pairs$citation[i], "-", pairs$license_url[i], "\n\n")
  }
}
```

---

Collection dates are provider-reported and may be unknown; publication or filename dates are never substituted. Downloads preserve original tiles, including portions outside the study area. This report is generated locally and is not transmitted anywhere.
