---
title: "ALS Downloader | Study report"
output:
  html_document:
    self_contained: true
params:
  tiles: NULL
  aoi_area_km2: NA
  aoi: NULL
  figures: NULL
  details: false
  software_citation: ""
geometry: margin=0.65in
fontsize: 10pt
header-includes:
  - \usepackage{helvet}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage{xcolor}
  - \definecolor{alsaccent}{HTML}{176B68}
  - \makeatletter
  - \renewcommand\section{\@startsection{section}{1}{\z@}{-2ex plus -1ex minus -.2ex}{1ex plus .2ex}{\normalfont\large\bfseries\color{alsaccent}}}
  - \def\@maketitle{\noindent{\color{alsaccent}\Huge\bfseries\@title\par}\vspace{0.7em}}
  - \renewcommand\subsection[1]{\par\vspace{1.4em}{\large\bfseries\color{alsaccent}#1\par}\vspace{.3em}{\color{alsaccent}\hrule}\vspace{.7em}}
  - \makeatother
---

<style>
body{max-width:900px;margin:auto;padding:32px 32px 60px;font-family:Arial,Helvetica,sans-serif;color:#263a40;line-height:1.5}
h1.title{color:#176b68;font-size:32px;font-weight:700}
h2{color:#176b68;border-bottom:2px solid #176b68;padding-bottom:6px;margin-top:26px;font-size:21px}
table{width:100%;border-collapse:collapse;font-size:14px;margin:12px 0 18px;background:#f2f5f5}
td,th{padding:9px 12px;border-bottom:3px solid white;text-align:left}
img{max-width:100%} .report-note{background:#f2f5f5;padding:16px}
@media print{h2{break-after:avoid}img,table{break-inside:avoid}}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "asis")
tiles <- params$tiles
aoi_area_km2 <- params$aoi_area_km2
aoi <- params$aoi
has_tiles <- !is.null(tiles) && is.data.frame(tiles) && nrow(tiles) > 0
has_aoi <- inherits(aoi, "sf") && nrow(aoi) > 0
has_map <- has_tiles && has_aoi && inherits(tiles, "sf")
tiles_df <- if (has_tiles && inherits(tiles, "sf")) sf::st_drop_geometry(tiles) else tiles
bytes <- if (has_tiles) suppressWarnings(as.numeric(tiles_df$size_bytes)) else numeric()
known <- is.finite(bytes) & bytes > 0
total_gb <- sum(bytes[known]) / 1024^3
providers <- if (has_tiles) sort(unique(tiles_df$provider)) else character()
dates <- if (has_tiles) sort(unique(na.omit(c(as.character(tiles_df$acquired_start), as.character(tiles_df$acquired_end))))) else character()
dates <- dates[nzchar(dates)]
area_label <- if (length(aoi_area_km2) == 1 && is.finite(aoi_area_km2)) sprintf("%.4f km2", aoi_area_km2) else "Not recorded"
date_label <- if (length(dates)) paste(min(dates), "to", max(dates)) else "Not reported"
```

Discover, inspect and download airborne LiDAR. A visual summary of your selected data, with source information to support the next step in your study.

## Key information

```{r}
key <- data.frame(
  Information = c("Report date", "Study area", "Selected data", "Acquisition interval", "Known storage"),
  Value = c(format(Sys.time(), "%Y-%m-%d %H:%M"), area_label,
    sprintf("Tiles: %d | Providers: %d", if (has_tiles) nrow(tiles) else 0, length(providers)),
    date_label, sprintf("%.3f GiB | Files with unknown size: %d", total_gb, sum(!known))))
if (knitr::is_latex_output()) {
  cat("```{=latex}\n\\colorbox{black!4}{\\parbox{\\dimexpr\\linewidth-2\\fboxsep}{\n")
  panel <- knitr::kable(key, format = "latex", row.names = FALSE, col.names = NULL, vline = "")
  cat("\\renewcommand{\\arraystretch}{1.5}\n")
  cat(gsub("\\hline", "", panel, fixed = TRUE))
  cat("}}\n```\n\n")
} else print(knitr::kable(key, row.names = FALSE, col.names = c("", "")))
if (length(providers)) cat("\nProviders:", paste(providers, collapse = ", "), "\n\n")
if (!has_tiles) cat("No tiles in this selection.\n\n")
```

## Study area and selected coverage

```{r fig.width=7.2, fig.height=4.1, fig.align="center"}
if (has_map) {
  aoi_geo <- sf::st_transform(aoi, 4326)
  tiles_geo <- sf::st_transform(tiles, 4326)
  reported <- ifelse(is.na(tiles_geo$acquired_end), tiles_geo$acquired_start, tiles_geo$acquired_end)
  year <- suppressWarnings(as.integer(substr(reported, 1, 4)))
  years <- sort(unique(stats::na.omit(year)))
  n_year <- length(years)
  cols <- grDevices::hcl.colors(max(n_year, 2), "Set 2")
  fill <- if (n_year > 0) cols[match(year, years)] else rep(NA_character_, nrow(tiles_geo))
  fill[is.na(fill)] <- "#8a97a1"
  combined <- c(sf::st_geometry(aoi_geo), sf::st_geometry(tiles_geo))
  bb <- sf::st_bbox(combined)
  graphics::par(mar = c(3.5, 5.5, 1, 1), family = "sans")
  plot(sf::st_geometry(tiles_geo), col = fill, border = "#1c2b35",
    xlim = bb[c("xmin", "xmax")], ylim = bb[c("ymin", "ymax")], main = "", axes = FALSE)
  graphics::axis(1, cex.axis = .8)
  graphics::axis(2, las = 1, cex.axis = .8)
  graphics::mtext("Longitude (WGS84)", side = 1, line = 2.2, cex = .8)
  graphics::mtext("Latitude (WGS84)", side = 2, line = 4.2, cex = .8)
  graphics::box()
  plot(sf::st_geometry(aoi_geo), add = TRUE, border = "#dc2626", lwd = 2, col = NA)
  if (n_year > 0) legend("topright", legend = c(as.character(years), "Unknown"),
    fill = c(cols[seq_len(n_year)], "#8a97a1"), bty = "n", cex = .8, title = "Acquisition year")
  cat("\n\n*Tile footprints (fill = acquisition year) and the study area outline (red). Geometry only, not a basemap image.*\n\n")
} else if (has_tiles) {
  cat("Study area geometry was not supplied to this report; the map is omitted.\n\n")
}
```

```{r attached-figures, out.width="100%", out.height="7in", fig.align="center"}
if (length(params$figures)) {
  for (i in seq_along(params$figures)) {
    if (knitr::is_latex_output()) cat("\\newpage\n")
    cat(sprintf("\n\n## Figure %d | Exported view\n\n", i))
    cat("User-selected exported view; its embedded labels, sampling notes and credits apply. It may differ from the current tile selection. Visual comparisons do not measure change.\n\n")
    if (knitr::is_latex_output()) cat(sprintf("\\includegraphics[width=\\linewidth,height=0.72\\textheight,keepaspectratio]{%s}\n", params$figures[i]))
    else cat(sprintf("![](%s)\n\n", params$figures[i]))
  }
}
```

```{r}
if (knitr::is_latex_output()) cat("\\newpage\n")
```

## Conclusions and next steps

```{r}
if (has_tiles) {
  cat(sprintf("- Selected tiles: **%d**. Data providers: **%d**. ", nrow(tiles), length(providers)))
  cat(if (length(dates)) paste0("Reported acquisition dates span **", date_label, "**; this is not a continuous coverage record.\n") else "Acquisition dates are not reported.\n")
  if (any(known)) cat(sprintf("- Plan for **%.3f GiB** of known source files%s. Allow additional space for temporary files and derived products.\n", total_gb,
    if (any(!known)) sprintf(" plus files with unknown sizes (%d)", sum(!known)) else ""))
  else cat("- No usable file sizes: storage and transfer time cannot be estimated.\n")
} else cat("No data conclusions can be drawn from an empty selection.\n")
```

Figures support visual inspection. They do not establish measured change, point-cloud alignment or data quality. Confirm acquisition dates, coordinate references and elevation units before quantitative analysis. Downloads preserve complete source tiles, including points outside the study area.

## Software citation

```{r}
cat(params$software_citation, "\n\n")
```

## Acknowledgements

ALS Downloader was developed within [OpenForest4D](https://openforest4d.org). OpenForest4D is funded by NSF awards **2409885, 2409886 & 2409887**.

We acknowledge the data producers and repositories listed below. Cite their datasets separately when using downloaded data or figures.

## Citations and licences

```{r}
if (has_tiles) {
  cols <- intersect(c("citation", "license_url"), names(tiles_df))
  if (length(cols) == 2) {
    pairs <- unique(tiles_df[, cols, drop = FALSE])
    for (i in seq_len(nrow(pairs))) {
      credit <- if (is.na(pairs$citation[i]) || !nzchar(pairs$citation[i])) "Source credit not supplied" else pairs$citation[i]
      link <- pairs$license_url[i]
      terms <- if (!is.na(link) && grepl("^https?://", link)) paste0("[Licence and source terms](<", link, ">)") else "Licence link not supplied"
      cat("-", credit, "-", terms, "\n\n")
    }
  }
}
```


Source datasets retain their own licences and access conditions. This report describes the current selection, not a completed download or a certification of coverage. Reports are generated by the app server; local mode keeps them on this computer.

```{r}
if (isTRUE(params$details)) {
  if (knitr::is_latex_output()) cat("\\newpage\n")
  cat(knitr::knit_child(system.file("report", "technical-appendix.Rmd", package = "alsdownloader"), quiet = TRUE), sep = "\n")
}
```
