---
title: "ALS Downloader session report"
output:
  html_document:
    self_contained: true
params:
  tiles: NULL
  aoi_area_km2: NA
  aoi: NULL
  figures: NULL
geometry: margin=0.7in
fontsize: 10pt
---

<style>
body{max-width:900px;margin:0 auto;padding:0 24px 70px;font-family:-apple-system,"Segoe UI",Helvetica,Arial,sans-serif;color:#222}
h1.title{color:#1c4a5e}
h2{border-bottom:1px solid #e2e2e2;padding-bottom:4px;margin-top:34px}
blockquote{color:#555;font-size:15px;font-style:italic;border-left:3px solid #8fb4c9;padding-left:12px;margin:10px 0 22px}
.als-footer-badge{position:fixed;right:18px;bottom:14px;background:#152936;color:#dbe7ee;font-size:11.5px;padding:5px 12px;border-radius:14px;letter-spacing:.3px;box-shadow:0 2px 8px rgba(0,0,0,.25)}
table{font-size:13.5px}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "asis")
tiles <- params$tiles
aoi_area_km2 <- params$aoi_area_km2
aoi <- params$aoi
has_tiles <- !is.null(tiles) && is.data.frame(tiles) && nrow(tiles) > 0
has_aoi <- inherits(aoi, "sf") && nrow(aoi) > 0
has_map <- has_tiles && has_aoi && inherits(tiles, "sf")
# The map figure needs geometry; every table below only ever needs the
# plain attribute columns, so it uses this geometry-free copy instead.
tiles_df <- if (has_tiles && inherits(tiles, "sf")) sf::st_drop_geometry(tiles) else tiles
```

Discover, inspect and download airborne LiDAR.

Generated `r format(Sys.time(), "%Y-%m-%d %H:%M")` by [ALS Downloader](https://github.com/Cesarito2021/als_downloader). Selected source metadata; not a completed-download certificate.

## Study area

```{r}
if (length(aoi_area_km2) == 1 && is.finite(aoi_area_km2)) {
  cat(sprintf("Approximately **%.4f km2**.\n\n", aoi_area_km2))
} else {
  cat("Not recorded.\n\n")
}
```

```{r fig.width=7.2, fig.height=3.4, fig.align="center"}
if (has_map) {
  aoi_geo <- sf::st_transform(aoi, 4326)
  tiles_geo <- sf::st_transform(tiles, 4326)
  reported <- ifelse(is.na(tiles_geo$acquired_end), tiles_geo$acquired_start, tiles_geo$acquired_end)
  year <- suppressWarnings(as.integer(substr(reported, 1, 4)))
  years <- sort(unique(stats::na.omit(year)))
  n_year <- length(years)
  cols <- grDevices::hcl.colors(max(n_year, 2), "Set 2")
  fill <- if (n_year > 0) cols[match(year, years)] else rep(NA_character_, nrow(tiles_geo))
  fill[is.na(fill)] <- "#8a97a1"
  combined <- c(sf::st_geometry(aoi_geo), sf::st_geometry(tiles_geo))
  bb <- sf::st_bbox(combined)
  plot(sf::st_geometry(tiles_geo), col = fill, border = "#1c2b35",
    xlim = bb[c("xmin", "xmax")], ylim = bb[c("ymin", "ymax")], main = "")
  plot(sf::st_geometry(aoi_geo), add = TRUE, border = "#dc2626", lwd = 2, col = NA)
  if (n_year > 0) legend("topright", legend = c(as.character(years), "Unknown"),
    fill = c(cols[seq_len(n_year)], "#8a97a1"), bty = "n", cex = .8, title = "Acquisition year")
  cat("\n\n*Tile footprints (fill = acquisition year) and the study area outline (red). Geometry only, not a basemap image.*\n\n")
} else if (has_tiles) {
  cat("Study area geometry was not supplied to this report; only the tile footprints below are known.\n\n")
}
```

## Summary

```{r}
if (!has_tiles) {
  cat("No tiles in this selection.\n")
} else {
  bytes <- suppressWarnings(as.numeric(tiles_df$size_bytes))
  known <- is.finite(bytes) & bytes > 0
  total_gb <- sum(bytes[known]) / 1024^3
  providers <- sort(unique(tiles_df$provider))
  cat(sprintf("- **%d** selected tile(s), **%d** provider(s)\n", nrow(tiles_df), length(providers)))
  cat(sprintf("- Providers: %s\n", paste(providers, collapse = ", ")))
  cat(sprintf("- **%.3f GiB** known source-file storage; **%d** file(s) have unknown size\n", total_gb, sum(!known)))
  if (any(!known)) cat("- Storage is incomplete: unknown files are excluded, not treated as empty.\n")
  dates <- sort(unique(na.omit(c(as.character(tiles_df$acquired_start), as.character(tiles_df$acquired_end)))))
  dates <- dates[nzchar(dates)]
  if (length(dates)) cat(sprintf("- Reported acquisition dates: **%s to %s** (interval endpoints, not continuous coverage)\n", min(dates), max(dates)))
  else cat("- Acquisition dates: not reported\n")

}
```

```{r}
if (knitr::is_latex_output()) cat("\\newpage\n")
```

## Download planning

Storage covers original compressed source tiles, including points outside the study area. Allow additional space for temporary downloads, extraction, decompressed LAS and derived products; their sizes are not estimated here. 1 GiB = 1,024 MiB.

```{r}
if (has_tiles && any(known)) {
  speed <- c(5, 20, 50)
  minutes <- sum(bytes[known]) / 1024^2 / speed / 60
  scenario <- data.frame(`Aggregate transfer (MiB/s)` = speed,
    `Time for known bytes (minutes)` = ifelse(minutes < 0.1, "<0.1", sprintf("%.1f", minutes)), check.names = FALSE)
  print(knitr::kable(scenario, row.names = FALSE))
} else cat("No usable file sizes: storage and transfer time cannot be estimated.\n\n")
```

Illustrative ideal transfer times, **not measured speeds or completion promises**. Unknown sizes, retries, throttling, disk writes and processing are excluded. Four workers share the connection and provider limits; four CPU cores do not imply four times the download speed.

## Selection sample (first 5 files)

```{r}
if (has_tiles) {
  cols <- intersect(c("filename", "dataset", "provider", "acquired_start", "acquired_end", "size_bytes"), names(tiles_df))
  show <- utils::head(tiles_df[, cols, drop = FALSE], 5L)
  # Keep long filenames out of the compact printed table.
  cols <- intersect(c("filename", "acquired_start", "acquired_end", "size_bytes"), names(show))
  show <- show[, cols, drop = FALSE]
  show$filename <- ifelse(nchar(show$filename) > 28, paste0(substr(show$filename, 1, 25), "..."), show$filename)
  if ("size_bytes" %in% names(show)) show$size_bytes <- round(show$size_bytes / 1024^2, 2)
  names(show) <- c(filename="File", acquired_start="Start", acquired_end="End", size_bytes="MiB")[cols]
  print(knitr::kable(show, row.names = FALSE))
}
```

Export selected metadata in the app for full filenames and the complete selection. Dates are provider-reported; publication dates are never substituted.

## Citations and licences

```{r}
if (has_tiles) {
  cols <- intersect(c("citation", "license_url"), names(tiles_df))
  if (length(cols) == 2) {
    pairs <- unique(tiles_df[, cols, drop = FALSE])
    for (i in seq_len(nrow(pairs))) {
      credit <- if (is.na(pairs$citation[i]) || !nzchar(pairs$citation[i])) "Source credit not supplied" else pairs$citation[i]
      link <- pairs$license_url[i]
      terms <- if (!is.na(link) && grepl("^https?://", link)) paste0("[Licence and source terms](<", link, ">)") else "Licence link not supplied"
      cat("-", credit, "-", terms, "\n\n")
    }
  }
}
```

Source datasets retain their own terms. This report does not certify data quality, complete coverage or licence clearance. Reports are generated by the app server; local mode keeps them on this computer.

```{r attached-figures, out.width="100%", out.height="7in", fig.align="center"}
if (length(params$figures)) {
  for (i in seq_along(params$figures)) {
    if (knitr::is_latex_output()) cat("\\newpage\n")
    cat(sprintf("\n\n## Attached figure %d\n\n", i))
    cat("User-selected exported view; its embedded labels, sampling notes and credits apply. It may differ from the current tile selection. Visual comparisons do not measure change.\n\n")
    if (knitr::is_latex_output()) cat(sprintf("\\includegraphics[width=\\linewidth,height=0.75\\textheight,keepaspectratio]{%s}\n", params$figures[i]))
    else cat(sprintf("![](%s)\n\n", params$figures[i]))
  }
}
```
