---
title: "ALS Downloader Report"
output:
  html_document:
    self_contained: true
params:
  tiles: NULL
  aoi_area_km2: NA
  aoi: NULL
  figures: NULL
  figure_captions: NULL
  map_image: NULL
  map_credits: ""
  details: false
  software_citation: ""
geometry: margin=0.65in
fontsize: 10pt
header-includes:
  - \usepackage{helvet}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage{xcolor}
  - \definecolor{alsaccent}{HTML}{176B68}
  - \makeatletter
  - \renewcommand\section{\@startsection{section}{1}{\z@}{-2ex plus -1ex minus -.2ex}{1ex plus .2ex}{\normalfont\large\bfseries\color{alsaccent}}}
  - \def\@maketitle{}
  - \renewcommand\subsection[1]{\par\vspace{1.4em}{\large\bfseries\color{alsaccent}#1\par}\vspace{.3em}{\color{alsaccent}\hrule}\vspace{.7em}}
  - \makeatother
---

<style>
body{max-width:900px;margin:auto;padding:32px 32px 60px;font-family:Arial,Helvetica,sans-serif;color:#263a40;line-height:1.5}
h1.title{color:#176b68;font-size:32px;font-weight:700}
h2{color:#176b68;border-bottom:2px solid #176b68;padding-bottom:6px;margin-top:26px;font-size:21px}
table{width:100%;border-collapse:collapse;font-size:14px;margin:12px 0 18px;background:#f2f5f5}
td,th{padding:9px 12px;border-bottom:3px solid white;text-align:left}
img{max-width:100%} .report-note{background:#f2f5f5;padding:16px}
@media print{h2{break-after:avoid}img,table{break-inside:avoid}}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "asis")
tiles <- params$tiles
aoi_area_km2 <- params$aoi_area_km2
aoi <- params$aoi
has_tiles <- !is.null(tiles) && is.data.frame(tiles) && nrow(tiles) > 0
has_aoi <- inherits(aoi, "sf") && nrow(aoi) > 0
has_map <- has_tiles && has_aoi && inherits(tiles, "sf")
tiles_df <- if (has_tiles && inherits(tiles, "sf")) sf::st_drop_geometry(tiles) else tiles
bytes <- if (has_tiles) suppressWarnings(as.numeric(tiles_df$size_bytes)) else numeric()
known <- is.finite(bytes) & bytes > 0
total_gb <- sum(bytes[known]) / 1024^3
providers <- if (has_tiles) sort(unique(tiles_df$provider)) else character()
dates <- if (has_tiles) sort(unique(na.omit(c(as.character(tiles_df$acquired_start), as.character(tiles_df$acquired_end))))) else character()
dates <- dates[nzchar(dates)]
area_label <- if (length(aoi_area_km2) == 1 && is.finite(aoi_area_km2)) sprintf("%.4f km2", aoi_area_km2) else "Not recorded"
date_label <- alsdownloader:::report_date_label(tiles_df$acquired_start, tiles_df$acquired_end)
storage_label <- alsdownloader:::report_storage_label(bytes)
has_map_figure <- length(params$map_image) > 0 || has_map
figure_number <- if (has_map_figure) 1L else 0L
```

```{r report-banner}
if (knitr::is_latex_output()) {
  banner <- normalizePath(system.file("report", "banner.png", package = "alsdownloader"), winslash = "/", mustWork = TRUE)
  cat(paste0("```{=latex}\n\\noindent\\begin{minipage}[c]{0.50\\linewidth}\n",
    "\\includegraphics[width=\\linewidth,height=2cm,keepaspectratio]{\\detokenize{", banner, "}}\n",
    "\\end{minipage}\\hfill\\begin{minipage}[c]{0.45\\linewidth}\\raggedleft\n",
    "{\\color{alsaccent}\\Large\\bfseries ALS Downloader Report}\\par\n",
    "\\end{minipage}\\par\\vspace{0.8em}\n```\n\n"))
}
```

Discover, inspect and download airborne LiDAR. A summary of the selected tiles, acquisition dates, storage and data sources.

## Key information

```{r}
key <- data.frame(
  Information = c("Report date", "Area of interest (AOI)", "Selected data", "Acquisition date(s)", "Storage"),
  Value = c(format(Sys.time(), "%Y-%m-%d %H:%M"), area_label,
    sprintf("Tiles: %d | Providers: %d", if (has_tiles) nrow(tiles) else 0, length(providers)),
    date_label, storage_label))
if (knitr::is_latex_output()) {
  cat("```{=latex}\n\\colorbox{black!4}{\\parbox{\\dimexpr\\linewidth-2\\fboxsep}{\n")
  panel <- knitr::kable(key, format = "latex", row.names = FALSE, col.names = NULL, vline = "")
  cat("\\renewcommand{\\arraystretch}{1.5}\n")
  cat(gsub("\\hline", "", panel, fixed = TRUE))
  cat("}}\n```\n\n")
} else print(knitr::kable(key, row.names = FALSE, col.names = c("", "")))
if (length(providers)) cat("\nProviders:", paste(providers, collapse = ", "), "\n\n")
if (!has_tiles) cat("No tiles in this selection.\n\n")
```

```{r fig.width=7.2, fig.height=4.1, fig.align="center"}
if (length(params$map_image)) {
  if (knitr::is_latex_output()) cat(sprintf("\\begin{center}\n\\includegraphics[width=\\linewidth,height=0.45\\textheight,keepaspectratio]{%s}\n\\end{center}\n\n", params$map_image))
  else cat(sprintf("![](%s)\n\n", params$map_image))
  cat("**Figure 1.** Study area (gold outline) and selected tile footprints.\n\n")
  cat(params$map_credits, "\n\n")
} else if (has_map) {
  aoi_geo <- sf::st_transform(aoi, 4326)
  tiles_geo <- sf::st_transform(tiles, 4326)
  reported <- ifelse(is.na(tiles_geo$acquired_end), tiles_geo$acquired_start, tiles_geo$acquired_end)
  year <- suppressWarnings(as.integer(substr(reported, 1, 4)))
  years <- sort(unique(stats::na.omit(year)))
  n_year <- length(years)
  cols <- grDevices::hcl.colors(max(n_year, 2), "Set 2")
  fill <- if (n_year > 0) cols[match(year, years)] else rep(NA_character_, nrow(tiles_geo))
  fill[is.na(fill)] <- "#8a97a1"
  combined <- c(sf::st_geometry(aoi_geo), sf::st_geometry(tiles_geo))
  bb <- sf::st_bbox(combined)
  graphics::par(mar = c(3.5, 5.5, 1, 1), family = "sans")
  plot(sf::st_geometry(tiles_geo), col = fill, border = "#1c2b35",
    xlim = bb[c("xmin", "xmax")], ylim = bb[c("ymin", "ymax")], main = "", axes = FALSE)
  graphics::axis(1, cex.axis = .8)
  graphics::axis(2, las = 1, cex.axis = .8)
  graphics::mtext("Longitude (WGS84)", side = 1, line = 2.2, cex = .8)
  graphics::mtext("Latitude (WGS84)", side = 2, line = 4.2, cex = .8)
  graphics::box()
  plot(sf::st_geometry(aoi_geo), add = TRUE, border = "#dc2626", lwd = 2, col = NA)
  if (n_year > 0) legend("topright", legend = c(as.character(years), "Unknown"),
    fill = c(cols[seq_len(n_year)], "#8a97a1"), bty = "n", cex = .8, title = "Acquisition year")
  cat("\n\n**Figure 1.** Study area (red outline) and selected tile footprints, coloured by acquisition year.\n\n")
} else if (has_tiles) {
  # No map section when no map was generated.
}
```

```{r attached-figures, out.width="100%", out.height="7in", fig.align="center"}
if (length(params$figures)) {
  for (i in seq_along(params$figures)) {
    if (knitr::is_latex_output()) cat("\\newpage\n")
    figure_number <- figure_number + 1L
    if (knitr::is_latex_output()) cat(sprintf("\\includegraphics[width=\\linewidth,height=0.72\\textheight,keepaspectratio]{%s}\n", params$figures[i]))
    else cat(sprintf("![](%s)\n\n", params$figures[i]))
    cat(sprintf("\n\n**Figure %d.** %s\n\n", figure_number, params$figure_captions[i]))
  }
}
```

```{r}
if (knitr::is_latex_output()) cat("\\newpage\n")
```

## Summary

```{r}
if (has_tiles) {
  cat(sprintf("Selected tiles: **%d**. Data providers: **%d**.\n\n", nrow(tiles), length(providers)))
  cat("Acquisition date(s): **", date_label, "**.\n\n", sep="")
  cat("Storage: **", storage_label, "**.\n\n", sep="")
} else cat("No tiles in this selection.\n\n")
```

Downloaded point clouds are original files from their providers. ALS Downloader applies no additional pre- or post-processing to these downloads. Figures show sampled point clouds for visual inspection.

## Data source licences and policies

```{r}
if (has_tiles) {
  cols <- intersect(c("citation", "license_url"), names(tiles_df))
  if (length(cols) == 2) {
    pairs <- unique(tiles_df[, cols, drop = FALSE])
    for (i in seq_len(nrow(pairs))) {
      credit <- if (is.na(pairs$citation[i]) || !nzchar(pairs$citation[i])) "Source credit not supplied" else pairs$citation[i]
      link <- pairs$license_url[i]
      terms <- if (!is.na(link) && grepl("^https?://", link)) paste0("[Licence and source terms](<", link, ">)") else "Licence link not supplied"
      cat("-", credit, "-", terms, "\n\n")
    }
  }
}
```


Source datasets retain their own licences. Cite the original data producers and dataset DOIs when using these files or figures.

```{r}
if (isTRUE(params$details)) {
  if (knitr::is_latex_output()) cat("\\newpage\n")
  cat(knitr::knit_child(system.file("report", "technical-appendix.Rmd", package = "alsdownloader"), quiet = TRUE), sep = "\n")
}
```

## Software citation

```{r}
cat(params$software_citation, "\n\n")
```

## Acknowledgements

The development of ALS Downloader at the **University of Florida**, within [OpenForest4D](https://openforest4d.org), was supported by the **U.S. National Science Foundation** under awards **2409885, 2409886 and 2409887**. We thank the data producers and repositories cited in this report.

Any opinions, findings and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
