## Technical appendix

### Download planning

Storage covers original compressed source tiles, including points outside the AOI. Allow additional space for temporary downloads, extraction, decompressed LAS and derived products; their sizes are not estimated here. 1 GiB = 1,024 MiB.

```{r, eval=params$details}
if (has_tiles && any(known)) {
  speed <- c(5, 20, 50)
  minutes <- sum(bytes[known]) / 1024^2 / speed / 60
  scenario <- data.frame(`Aggregate transfer (MiB/s)` = speed,
    `Time for known bytes (minutes)` = ifelse(minutes < 0.1, "<0.1", sprintf("%.1f", minutes)), check.names = FALSE)
  print(knitr::kable(scenario, row.names = FALSE))
} else cat("No usable file sizes: storage and transfer time cannot be estimated.\n\n")
```

Illustrative ideal transfer times, **not measured speeds or completion promises**. Unknown sizes, retries, throttling, disk writes and processing are excluded. Four workers share the connection and provider limits; four CPU cores do not imply four times the download speed.

### Metadata and source evidence

Export selected metadata for file identifiers, acquisition evidence and source links. Per-file sizes are retained internally; the report presents aggregate download size. NA marks unavailable information, with lookup status retained separately.

```{r, eval=params$details}
if(has_tiles) {
  fields<-intersect(c("filename","acquisition_year","date_status"),names(tiles_df))
  print(knitr::kable(utils::head(tiles_df[fields],5),row.names=FALSE))
}
```
