---
title: "Discovering and downloading ALS tiles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Discovering and downloading ALS tiles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The package exposes the same study-area and transfer functions used by its Shiny
interface. Package loading and the examples below make no network requests.

```{r}
library(ALSdownloadeR)
ring <- matrix(c(0,0, 1,0, 1,1, 0,1, 0,0), ncol=2, byrow=TRUE)
aoi <- read_aoi(sf::st_sfc(sf::st_polygon(list(ring)), crs=4326))
aoi_area(aoi)
download_worker_policy("local", available_cores=8)
```

## Interactive use

Start the app explicitly from an interactive R session:

```{r, eval=FALSE}
launch_app()
```

Draw an area or upload a zipped Shapefile, GeoPackage or GeoJSON. Select a
provider and acquisition interval, search, then select tile rows. For local
downloads choose an output directory. Unavailable acquisition years remain visible as NA.

## Scripted use

These commands need a real study area and internet access and are not executed
when this vignette is built:

```{r, eval=FALSE}
aoi <- read_aoi("study-area.gpkg", layer="study")
tiles <- find_als_data(aoi, provider="usgs3dep")
summarize_als_download(tiles)
download_als_data(tiles, output_dir="selected-tiles", workers=2)

ot_tiles <- find_als_data(aoi, provider="opentopography")
```

OpenTopography uses the verified hosted airborne-LiDAR catalogue by default.
Matching original tile indexes load automatically; acquisition dates, citations
and licence links come from verified collection metadata. Missing dates remain
NA. An optional `tile_index_dir` supports legacy, locally supplied indexes.
Exact overview outlines use a checksum-pinned auxiliary catalogue downloaded
only when the interactive map needs detailed coverage; the small packaged search
index suffices to locate candidate collections. Final AOI selection always uses
original provider tile polygons.
USGS 3DEP searches use the official TNM inventory and original files. The optional
Planetary Computer mirror signs its Azure asset links immediately before transfer.
The final acquisition year is displayed with its evidence; filename references
are labelled as unverified. Viewing remote point clouds and comparisons downloads
the complete selected files temporarily before reading and plotting them.

## Resume and validation

Each completed tile has a sidecar record with its source and checksum. Running
the same request again skips only files whose records and checksums match.
Inspect `manifest.csv` for failures. HTTP status, reported length and the LAS
signature are checked; complete decompression of every point is not verified.
Existing untracked destination files are not overwritten.

Transfer failures remain in the result table. Rate-limit and authorization
responses stop that tile rather than causing aggressive automatic retries.
After a hard crash, verify that no transfer process is alive before removing
the output directory's `.als-transfer-lock` directory.

## Preview and units

```{r}
p <- preview_points(data.frame(X=1:100, Y=101:200, Z=201:300), max_points=12)
nrow(p)
attr(p, "origin")
```

`read_preview()` optionally uses lidR to decimate a local LAS/LAZ file while
reading. It is not a remote COPC streaming implementation. Preview coordinates
retain the source units; the source's vertical datum and units must be checked.
Elevation colors do not represent terrain-normalized canopy height.

## Hosted deployment

Hosted mode uses serial downloads with a small per-batch storage allowance.
The administrator configures the OpenTopography index path and any shared lock
directory. The browser cannot select arbitrary server directories. Outputs are
temporary and must be saved before closing the session. Multi-process hosting
requires a shared `ALS_HOST_LOCK_DIR`; multiple hosts require a distributed job
queue, which this version does not provide.
